---
title: "Exam2"
author: "Kurata"
date: "6/27/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Q1 & 2 
```{r include = FALSE}
#clear environment
rm(list=ls(all=TRUE))

#load inequality data
library(rio)
inequality_data = import("inequality.xlsx")
```

### Q3
This is a cross-sectional data set as the summary for year below suggests that all data is from 2015, and the inequality_gini variable is recorded by country. 
```{r include=TRUE}
head(inequality_data)
str(inequality_data)
summary(inequality_data$year)
```
### Q4
```{r echo = FALSE}
subset(inequality_data, country == "Sweden")
subset(inequality_data, country == "Denmark")

```

### Q5
```{r echo = FALSE}
subset(inequality_data, country == "Brazil")
```

### Q6
It is better to have lower inequality_gini scores. 

### Q7
```{r echo= FALSE}
head(inequality_data)
```

### Q8 
```{r include = FALSE}
accent.remove = function(s){
  old1 = "ú"
  new1 = "u"
  s1 = chartr(old1, new1, s)
}

#apply function to remove accents 
inequality_data$country = accent.remove(inequality_data$country)

```
```{r echo = FALSE}
head(inequality_data)
```
### Q9
```{r include=FALSE}
inequality_data = inequality_data[order(inequality_data$inequality_gini),]
```
```{r echo = FALSE}
head(inequality_data,n=5)
```
### Q10
```{r include = TRUE}
avg = mean(inequality_data$inequality_gini, na.rm = TRUE)
```
### Q11 create new variables 
```{r include = FALSE}
inequality_data$low_inequality = NA
inequality_data$high_inequality = NA

inequality_data$low_inequality = ifelse(test = inequality_data$inequality_gini < avg,1,0)
inequality_data$high_inequality = ifelse(test = inequality_data$inequality_gini > avg,1,0)
```
### Q12 crosstab 
```{r echo = FALSE}
library(doBy)
summaryBy(low_inequality ~ high_inequality, data=inequality_data)
```
### Q13 for loop
```{r echo = FALSE}
actors = c("World Bank", "African Development Bank", "Bill and Melinda Gates Foundation")
for (i in 1:length(actors)){
  print(actors[i])
}
```
### Q14
Government Expenditure on Education may correlate to inequality, since I assume less equal countries to have less accessible education. 

### Q15 import variables
```{r include = FALSE}
library(WDI)
education_expenditure = WDI(country = "all",
                      indicator = c("SE.XPD.TOTL.GB.ZS"),
                      start = 2015, end = 2015, extra = FALSE, cache = NULL)
```
### Q16 rename variable 
```{r}
library(data.table)
setnames(education_expenditure,"SE.XPD.TOTL.GB.ZS","total_education_expenditure")
```

### Q17 merge data sets 
```{r include = FALSE}
library(tidyverse)
merged_df = left_join(x = inequality_data, 
                        y = education_expenditure,
                        by = "country")

#remove .x and .y variables and renaming them
merged_df = 
  merged_df %>%
  select(-c("year.x")) %>%
  dplyr::rename("year" = "year.y")

merged_df = 
  merged_df %>%
  select(-c("iso2c.x")) %>%
  dplyr::rename("iso2c" = "iso2c.y")
```
### Q18 remove countries with NA
```{r include = FALSE}
merged_df = na.omit(merged_df, select = c("inequality_gini","total_education_expenditure"))
```
### Q19 filter data 
```{r include = FALSE}
data_greater_30 =
  merged_df %>%
  dplyr::filter((inequality_gini>30))
```
### Q20 count
```{r}

```
### Q21 apply
```{r echo = FALSE}
sapply(data_greater_30$inequality_gini,sum)
```
### Q22 label 
```{r include = FALSE}
library(labelled)
var_label(merged_df) = list(`country` = "country",
                             `inequality_gini` = "distribution of income",
                             `low_inequality` = "below average inequality_gini",
                             `high_inequality` = "above average inequality_gini",
                             `iso2c` = "ISO-2 country code",
                             `total_education_expenditure` = "total government education expenditure",
                             `year` = "year")
```
### Q23 save labels
```{r include = FALSE}
export(merged_df,
       "final_data.dta")
```



